---
title: "goodpractice"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{goodpractice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why goodpractice?

`R CMD check` catches missing documentation and undeclared dependencies, but it stops there.
It won't tell you that your functions are too complex, your tests don't cover much, your code style is inconsistent, or your DESCRIPTION is missing a BugReports field.

goodpractice bundles `R CMD check` with code coverage ([covr](https://cran.r-project.org/package=covr)), source linting ([lintr](https://cran.r-project.org/package=lintr)), cyclomatic complexity ([cyclocomp](https://cran.r-project.org/package=cyclocomp)), and its own checks into a single report.
Each message tells you _what_ to fix, _why_ it matters, and _where_ in your code to look.

## Quick start

Point `gp()` at a package and it tells you what to fix.

```{r}
library(goodpractice)

pkg_path <- system.file("bad1", package = "goodpractice")
```

```{r, echo = TRUE, eval = FALSE}
g <- gp(pkg_path)
```
```{r, echo = FALSE, eval = TRUE}
g <- suppressWarnings(gp(pkg_path))
```

```{r}
g
```

Each line is one failed check.
The cross marks a problem, the text explains _what_ to fix and _why_, and the file path shows _where_.
If everything passes, you get praise instead.

## What gp() actually does

When you call `gp()`, two things happen in sequence:

1. **Gather data** — goodpractice runs a set of _preparation steps_ ("preps").
   One prep runs `R CMD check`.
   Another computes code coverage.
   Another lints your source files.
   Each prep runs once and stores its results.

2. **Run checks** — each check reads the stored results and returns pass, fail, or skip.
   A single prep can feed dozens of checks.
   The `rcmdcheck` prep alone powers over 200.

This two-step design means you never run `R CMD check` twice, even if 200 checks need its output.
It also means you can skip an entire category by excluding its prep — more on that below.

## Choosing what to run

By default, `gp()` runs `default_checks()` — about 250 checks covering documentation, code style, test coverage, namespace hygiene, and CRAN compliance.

```{r}
length(default_checks())
```

You can narrow things down by passing specific check names:

```{r}
g_url <- gp(pkg_path, checks = "description_url")
g_url
```

Or widen them by adding the optional tidyverse style checks:

```{r, eval = FALSE}
g <- gp(".", checks = c(default_checks(), tidyverse_checks()))
```

Three helper functions give you the available check names:

- `default_checks()` — the standard set
- `tidyverse_checks()` — opt-in style checks for tidyverse conventions
- `all_checks()` — both combined

```{r}
length(tidyverse_checks())
length(all_checks())
```

## Skipping slow preps

The checks themselves are fast.
What takes time is data gathering — `covr` has to install your package and run tests, `rcmdcheck` runs the full `R CMD check` suite.

If you only care about code style right now, skip the slow preps:

```{r, eval = FALSE}
options(goodpractice.exclude_preps = c("covr", "rcmdcheck"))
gp(".")
```

Every check that depends on a skipped prep gets skipped too.

You can also set this via the `GP_EXCLUDE_PREPS` environment variable, which is useful in CI:

```bash
GP_EXCLUDE_PREPS=covr,rcmdcheck Rscript -e 'goodpractice::gp(".")'
```

The R option takes precedence when both are set.
Exclusion only applies to the default check set — if you explicitly pass check names via the `checks` argument, exclusion is ignored.

## Check groups

Every check belongs to a group identified by its prep name.
This is the string you pass to `goodpractice.exclude_preps`:

| Prep name | What it checks |
|-----------|----------------|
| `rcmdcheck` | ~200 checks from `R CMD check`: documentation, namespace, compilation, tests, vignettes, CRAN compliance |
| `covr` | Whether all code is covered by tests |
| `cyclocomp` | Whether functions are too complex (default limit: 50) |
| `description` | Package metadata: URL, BugReports, dependency fields, author roles |
| `lintr` | Assignment style, line length, semicolons, `sapply()`, `library()` in package code |
| `namespace` | Whether `importFrom()` is used instead of whole-package imports, no `exportPattern()` |
| `rd` | Whether exported functions have `\examples` and `\value` sections |
| `spelling` | Misspelled words in documentation |
| `urlchecker` | Broken or redirecting URLs in documentation |
| `available_packages` | Whether CRAN reverse dependencies exist (informational) |
| `tidyverse_lintr` | Tidyverse style linting (opt-in via `tidyverse_checks()`) |
| _(none)_ | README exists, NEWS exists, `.R` file extension, `print()` returns invisibly |

Checks with no prep run instantly and cannot be excluded by prep name.
Use the `checks` argument to select or deselect them individually.

## Tidyverse style checks

The default checks stay away from style preferences — they focus on things that are objectively good practice regardless of how you format your code.

If your team follows the [tidyverse style guide](https://style.tidyverse.org/), there is an optional set for that:

```{r, eval = FALSE}
gp(".", checks = c(default_checks(), tidyverse_checks()))
```

Most of these are powered by [lintr](https://cran.r-project.org/package=lintr) — brace placement, spacing, naming conventions.
If your package has a `.lintr` configuration file, those settings are respected.

A few go beyond linting: whether R file names use snake_case, whether test files mirror source files, whether functions use `missing()` where `NULL` defaults would be clearer, and whether exported functions appear before internal helpers.

```{r}
tidyverse_checks()
```

## Working with results

The object returned by `gp()` has a few accessor functions beyond printing:

```{r}
checks(g_url)
failed_checks(g)
results(g)[1:5, ]
```

`checks()` lists what ran.
`failed_checks()` lists what failed.
`results()` gives you a data frame of every check and whether it passed.

An `NA` in the results means the check was skipped — usually because its prep failed or was excluded.

You can also export results to JSON with `export_json()`, and control how many file positions are printed with `print(g, positions_limit = Inf)`.

## Custom checks

You can add your own checks — see `vignette("custom_checks")` for the full guide.
