---
title: "goodpractice"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{goodpractice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Why goodpractice?

A robust R package needs more than passing `R CMD check`.
Tests give you confidence that things work.
A consistent coding style makes your code readable to others.
Keeping functions short and focused reduces the risk of bugs.
Good metadata — a BugReports URL, properly declared dependencies, documented return values — helps users and reviewers understand what they're working with.

goodpractice checks all of this in one pass.
It bundles `R CMD check` with code coverage ([covr](https://cran.r-project.org/package=covr)), source linting ([lintr](https://cran.r-project.org/package=lintr)), cyclomatic complexity ([cyclocomp](https://cran.r-project.org/package=cyclocomp)), and its own checks for documentation, package structure, and common pitfalls.
Each message tells you _what_ to fix, _why_ it matters, and _where_ in your code to look.
You can also add [custom checks](custom_checks.html) for your team's own conventions.

## Quick start

The main function is `gp()` (short for `goodpractice()`).
Point it at a package directory and it runs every check it knows:

```{r}
library(goodpractice)

# goodpractice ships with an example package that has some issues
pkg_path <- system.file("bad1", package = "goodpractice")
```

```{r, echo = TRUE, eval = FALSE}
g <- gp(pkg_path)
```
```{r, echo = FALSE, eval = TRUE}
g <- suppressWarnings(gp(pkg_path))
```

Printing the result shows only the checks that failed:

```{r}
g
```

Each line starting with a cross is one failed check.
The text after the cross explains what to fix and why.
The indented file paths below it show exactly where in your code the problem was found — in terminals and IDEs that support it, these paths are clickable.
If every check passes, you get a short praise message instead.

## What gp() actually does

When you call `gp()`, two things happen in sequence:

1. **Gather data** — goodpractice runs a set of _preparation steps_ ("preps").
   One prep runs `R CMD check`.
   Another computes code coverage by installing your package and exercising the tests.
   Another lints your source files.
   Each prep runs exactly once and stores its results for the checks to use.

2. **Run checks** — each check reads from the stored prep results and returns pass, fail, or skip.
   A single prep can feed many checks — the `rcmdcheck` prep alone powers over 200 individual checks, all drawn from a single `R CMD check` run.

This two-step design is why goodpractice can run 250+ checks without repeating expensive work.
It also means you can skip an entire category of checks by excluding its prep — more on that in the "Skipping slow preps" section below.

## Choosing what to run

By default, `gp()` runs everything in `default_checks()` — about 250 checks covering documentation, code style, test coverage, namespace hygiene, and CRAN compliance:

```{r}
length(default_checks())
```

If you only need a specific check, pass its name to the `checks` argument.
You can find check names with `all_checks()` and filter with `grep()`:

```{r}
# find checks related to URLs
grep("url", all_checks(), value = TRUE)
```

Then run just the ones you care about:

```{r}
g_url <- gp(pkg_path, checks = "description_url")
g_url
```

To go the other direction and run _more_ than the defaults, combine check sets.
For example, to add the optional tidyverse style checks on top of the defaults:

```{r, eval = FALSE}
g <- gp(".", checks = c(default_checks(), tidyverse_checks()))
```

Three helper functions give you the available check names:

- `default_checks()` — the standard set, run by default
- `tidyverse_checks()` — opt-in style checks following [tidyverse conventions](https://style.tidyverse.org/)
- `all_checks()` — both combined

```{r}
length(tidyverse_checks())
length(all_checks())
```

## Skipping slow preps

The checks themselves run in milliseconds — what takes time is the data gathering.
Computing code coverage with `covr` requires installing your package and running all your tests.
Running `R CMD check` via `rcmdcheck` exercises documentation, examples, vignettes, and compilation.
On a large package, these two preps alone can take several minutes.

If you only care about code style or DESCRIPTION metadata right now, you can skip the slow preps entirely.
Set the `goodpractice.exclude_preps` option to a character vector of prep names you want to skip:

```{r, eval = FALSE}
# skip coverage and R CMD check — just run the fast checks
options(goodpractice.exclude_preps = c("covr", "rcmdcheck"))
gp(".")
```

Every check that depends on a skipped prep is automatically skipped too — it shows up as `NA` in the results rather than pass or fail.

For CI/CD pipelines, you can set this via the `GP_EXCLUDE_PREPS` environment variable instead of modifying R code:

```bash
GP_EXCLUDE_PREPS=covr,rcmdcheck Rscript -e 'goodpractice::gp(".")'
```

When both the R option and the environment variable are set, the R option wins.
Exclusion only applies when you use the default check set — if you explicitly pass check names to the `checks` argument, exclusion settings are ignored and those checks always run.

```{r, eval = FALSE}
options(goodpractice.exclude_preps = "covr")

# covr checks are excluded here (using defaults)
gp(".")

# but covr checks run here because we asked for them explicitly
gp(".", checks = "covr")
```

## Check groups

Every check belongs to a group defined by the prep that gathers its data.
The table below lists every group by its prep name — the exact string you pass to `goodpractice.exclude_preps` when you want to skip it:

| Prep name | What it checks |
|-----------|----------------|
| `rcmdcheck` | ~200 checks from `R CMD check`: documentation, namespace, compilation, tests, vignettes, CRAN compliance |
| `covr` | Whether all code is covered by tests |
| `cyclocomp` | Whether functions are too complex (default limit: 50) |
| `description` | Package metadata: URL, BugReports, dependency fields, author roles |
| `lintr` | Assignment style, line length, semicolons, `sapply()`, `library()` in package code |
| `namespace` | Whether `importFrom()` is used instead of whole-package imports, no `exportPattern()` |
| `rd` | Whether exported functions have `\examples` and `\value` sections |
| `spelling` | Misspelled words in documentation |
| `urlchecker` | Broken or redirecting URLs in documentation |
| `available_packages` | Whether CRAN reverse dependencies exist (informational) |
| `tidyverse_lintr` | Tidyverse style linting (opt-in via `tidyverse_checks()`) |
| _(none)_ | README exists, NEWS exists, `.R` file extension, `print()` returns invisibly |

Checks in the last row have no prep — they inspect the package directory directly and run instantly.
Because they don't belong to a prep, you cannot skip them with `goodpractice.exclude_preps`.
To exclude them, pass the check names you _do_ want to the `checks` argument instead.

## Tidyverse style checks

The default checks deliberately stay away from style preferences — they focus on things that are good practice regardless of how you format your code.

If your team follows the [tidyverse style guide](https://style.tidyverse.org/), you can opt into an additional set of 30 style checks:

```{r, eval = FALSE}
# add tidyverse checks to the defaults
gp(".", checks = c(default_checks(), tidyverse_checks()))

# or run only the tidyverse checks
gp(".", checks = tidyverse_checks())
```

Most of these are powered by [lintr](https://cran.r-project.org/package=lintr) — brace placement, spacing, naming conventions, and so on.
They run `lintr::lint_package()` once and share the results across all lintr-based checks.
If your package has a `.lintr` configuration file, those settings are respected: disabled linters stay disabled, exclusions are honoured.

A few tidyverse checks go beyond linting and look at package structure directly — whether R file names use snake_case, whether test files mirror source files, whether functions use `missing()` where `NULL` defaults would be clearer, and whether exported functions appear before internal helpers.

To see all available tidyverse check names:

```{r}
tidyverse_checks()
```

## Working with results

Beyond printing, the object returned by `gp()` gives you programmatic access to the results.

`checks()` returns the names of all checks that were run:

```{r}
checks(g_url)
```

`failed_checks()` returns only the names of the checks that failed:

```{r}
failed_checks(g)
```

`results()` gives you a data frame with one row per check and a `passed` column that is `TRUE`, `FALSE`, or `NA` (skipped):

```{r}
results(g)[1:5, ]
```

A check shows `NA` when its prep failed or was excluded — it was not evaluated, so it neither passed nor failed.

To see all file positions for failed checks (not just the first five that `print()` shows), use `print()` with a higher limit:

```{r, eval = FALSE}
print(g, positions_limit = Inf)
```

You can also export the full results to JSON for use in other tools:

```{r, eval = FALSE}
export_json(g, "gp_results.json")
```

## Custom checks

goodpractice is extensible — you can define your own checks and preparation steps to enforce team-specific conventions.
See `vignette("custom_checks")` for the full guide.
