---
title: "goodpractice"
author: "Hannah Frick"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{goodpractice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## What's it for?

Building an R package is a great way of encapsulating code, documentation and data in a single testable and easily distributable unit.

For a package to be distributed via CRAN, it needs to pass a set of checks implemented in `R CMD check`, such as: Is there minimal documentation, e.g., are all arguments of exported functions documented? Are all dependencies declared?

These checks are helpful in developing a solid R package but they don't check for several other good practices. For example, a package does not need to contain any tests but is it good practice to include such. Following a coding standard helps readability. Avoiding overly complex functions reduces the risk of bugs. Including an URL for bug reports lets people more easily report bugs if they find any. 

Tools for automatically checking several of these aspects already exist and the **goodpractice** package bundles the checks from [**rcmdcheck**](https://cran.r-project.org/package=rcmdcheck) with code coverage through the [**covr**](https://cran.r-project.org/package=covr) package, source code linting via the [**lintr**](https://cran.r-project.org/package=lintr) package and cyclomatic complexity via the [**cyclocomp**](https://cran.r-project.org/package=cyclocomp) package and augments it with some further checks on good practice for R package development such as avoiding `T` and `F` in favour of `TRUE` and `FALSE`. It provides advice on which practices to follow and which to avoid.

You can use **goodpractice** checks as a reminder for you and your colleagues - and if you have custom checks to run, you can make **goodpractice** run those as well! Please see the vignette "Custom Checks" for more details.


## Good practice out of the box

### Main function

The main function is `goodpractice()` and has an alias `gp()` which takes the path to the source code of a package as its first argument. The **goodpractice** package contains the source for a simple package which violates some good practices. We'll use this for the examples.

```{r}
library(goodpractice)

# get path to example package
pkg_path <- system.file("bad1", package = "goodpractice")
```

```{r, echo = TRUE, eval = FALSE}
# run gp() on it
g <- gp(pkg_path)
```
```{r, echo = FALSE, eval = TRUE}
g <- suppressWarnings(gp(pkg_path))
```

```{r}
# show the result
g
```

So with this package, we've done a few things in the DESCRIPTION file for which there are reasons not to do them, have unnecessary trailing semicolons in the code and used `T` and `F` for `TRUE` and `FALSE`. The output of `gp()` tells you what you did that isn't considered good practice and if it's in the R code, it points you the location of your faux-pas. In general, the messages are supposed to not only point out to you _what_ you might want to avoid but also _why_.


The above example tries to run all 230 checks available, to see the full list use `all_checks()`. If you only want to run a subset of the checks, e.g., the one on the URL field in the DESCRIPTION, you can specify the checks by name:

```{r}
# what is the name of the check?
grep("url", all_checks(), value = TRUE)

# run only this check
g_url <- gp(pkg_path, checks = "description_url")

g_url
```

### Doing more than just printing

Apart from printing a `goodPractice` object as returned by `gp()` to access the advice, you can also access which checks were carried out and which of those failed:

```{r}
# which checks were carried out?
checks(g_url)

# which checks failed?
failed_checks(g)
```

To access all the checks carried out with their results in a data frame, use `results()` on your `goodPractice` object.

```{r}
# show the first 5 checks carried out and their results
results(g)[1:5,]
```

Note that the code coverage could not be calculated. The corresponding check does not show up in the failed checks (because it was not carried out) and the result is `NA`. It is also possible to export the results to a JSON file with `export_json()`.


## Tidyverse style checks

The default set of checks focuses on universal good practices — things every
R package benefits from regardless of coding style. But if your team follows
the [tidyverse style guide](https://style.tidyverse.org/), there is an
optional set of checks for that too.

These checks are not included in `default_checks()`. You opt in explicitly:

```{r, eval = FALSE}
# run default checks plus tidyverse style checks
g <- gp(".", checks = c(default_checks(), tidyverse_checks()))

# or run only the tidyverse checks
g <- gp(".", checks = tidyverse_checks())
```

Most tidyverse checks are powered by [**lintr**](https://cran.r-project.org/package=lintr)
under the hood — brace placement, spacing, naming conventions, and so on.
They run `lintr::lint_package()` once and share the results across all
lintr-based checks. If your package already has a `.lintr` configuration
file, those settings are respected: disabled linters stay disabled,
exclusions are honoured.

A handful of tidyverse checks go beyond what lintr covers. These look at
package structure directly — whether R file names use snake_case, whether
test files mirror source files, whether functions rely on `missing()` instead
of `NULL` defaults, and whether exported functions appear before internal
helpers in each file.

```{r}
# see all available tidyverse checks
tidyverse_checks()
```


## Excluding check groups

Running all 230+ checks on a large package can take a while. The checks
themselves are fast — what takes time is the data gathering that happens
before them. Each check belongs to a _check group_ (internally called a
"prep"). The `lintr` group runs the linter once, `covr` computes code
coverage once, `rcmdcheck` runs `R CMD check` once, and every check in
that group draws from the same results.

So when `covr` alone takes three minutes and all you care about right now is
code style, why wait? Instead of picking individual checks by name, you can
exclude an entire check group — and every check that depends on it goes
with it. Use `all_check_groups()` to see the available group names, and
`checks_by_group()` to see which checks belong to each group.

### Using an R option

Set the `goodpractice.exclude_check_groups` option to a character vector of group names:

```{r, eval = FALSE}
# skip code coverage and cyclomatic complexity checks
options(goodpractice.exclude_check_groups = c("covr", "cyclocomp"))
gp(".")
```

### Using an environment variable

Set the `GP_EXCLUDE_CHECK_GROUPS` environment variable to a comma-separated list of group names:

```{r, eval = FALSE}
Sys.setenv(GP_EXCLUDE_CHECK_GROUPS = "covr,cyclocomp")
gp(".")
```

This is particularly useful in CI/CD pipelines where you may want to configure behaviour without modifying R code:

```bash
GP_EXCLUDE_CHECK_GROUPS=covr,rcmdcheck Rscript -e 'goodpractice::gp(".")'
```

### Precedence rules

- The R option (`goodpractice.exclude_check_groups`) takes precedence over the environment variable (`GP_EXCLUDE_CHECK_GROUPS`) when both are set.
- Exclusion only applies when `checks = NULL` (the default). If you explicitly pass check names to the `checks` argument, exclusion settings are ignored.

```{r, eval = FALSE}
options(goodpractice.exclude_check_groups = "covr")

# covr checks are excluded
gp(".")

# covr checks still run because checks are explicitly specified
gp(".", checks = "covr")
```
