---
title: "'goodpractice' for developers"
author: "Mark Padgham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{'goodpractice' for developers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The 'Makefile'

The 'goodpractice' repository includes a ['Makefile'](https://www.gnu.org/software/make/manual/make.html#Introduction).
This allows many commands to be run directly from a shell console (rather than within an R session).
The default behaviour is to list all options:

```{bash, eval = FALSE}
make
```


```{bash, echo = FALSE, output = TRUE, eval = !identical(.Platform$OS.type, "windows")}
cd ..
make help | sed -E 's/\x1b\[[0-9;]*m//g'
```

Any of those options can then be run as `make <command>`.
For example, this command checks for any issues with the `pkgdown` site:
```{bash, eval = FALSE}
make pkgdowncheck
```
```{r, echo = FALSE, eval = TRUE, output = TRUE}
cli::cli_alert_success("No problems found")
```

Note that some of the `Makefile` options depend on additional packages which may need to be installed.
These include:

- [`allcontributors`](https://docs.ropensci.org/allcontributors/)
- [`pkgcheck`](https://docs.ropensci.org/pkgcheck/)
- [`pkgdown`](https://pkgdown.r-lib.org/)

---

## Adding new checks

Checks are defined in groups.
Each group is generally defined within two files:
1. A `prep_<group>.R` file for collecting data needed for the check
2. A `chk_<group>.R` file defining the output structure of the check.

The `R/` directory here consistns almost entirely of paired files defining each check group:
```{bash, echo = FALSE, output = TRUE}
ls ../R
```

An example is the `cyclocomp` check, which is largely a wrapper around the external [`cyclocomp` package](https://github.com/gaborcsardi/cyclocomp).
The `prep_cyclocomp.R` file looks like this:
```{bash, echo = FALSE, output = TRUE}
cat ../R/prep_cyclocomp.R
```

That preparation steps runs the `cyclocomp::cyclocomp_package_dir()` function in the source directory of the package being checked.
The corresponding `check_cyclocomp.R` file then uses a variable called `state` that is accessible to all functions within the package.
The line above `PREPS$cyclocomp <-` creates an entry of `state$cyclocomp` containing the results of the preparation step that can then be accessed to defined the check output in `R/chk_cyclocomp.R`:

```{bash, echo = FALSE, output = TRUE}
cat ../R/chk_cyclocomp.R
```

All checks following this general pattern of defining `PREPS$<group> <- ` entries containing data from the check, then defining `CHECKS$<group>` entries for output format, using `state$<group>` to access the check data.
The best way to learn about check structure is to examine some of the existing files.
